The function-sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-08-25 crelle-dictionary.tex
Working through, fixing small problems

\begin{abstract}
We extend the function-sheaf dictionary to smooth commutative group schemes over finite fields. 
As an application, we explain how to geometrize and categorify quasicharacters of  algebraic tori, abelian varieties and unipotent K-wound groups over connected commutative algebraic groups over non-Archimedean local fields that admit a Neron model locally of finite type,
\end{abstract}

@@

We will describe the group $\QCiso{G}$ in Theorem~\ref{thm:snake}
and the sets $\Hom(\qcs{L},\qcs{L}')$ in Theorem~\ref{thm:autornaught}; in this way we provide a complete description of the category $\QC(G)$.
%
In the meantime, we close this section with an elementary observation about $\Hom(\qcs{L},\qcs{L}')$.

\begin{lemma}\label{lem:autornaught}
Let $G$ be a smooth commutative group scheme over $\Fq$.
If $\qcs{L}$ and $\qcs{L}'$ are quasicharacter sheaves on $G$ then
every $\rho\in \Hom(\qcs{L},\qcs{L}')$ is either trivial or an isomorphism. 
\end{lemma}

@@

Now suppose $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf on $G$.
Then $\bm^* \gqcs{L} \iso \gqcs{L} \boxtimes\gqcs{L} $ together with
\ref{QC.3} guarantee
that the function $\trFrob{(\gqcs{L},\phi)} : G(\Fq)\to \EEx$ is a group homomorphism, which we will also denote by $\trFrob{\qcs{L}}$.  Moreover,
this homomorphism depends only on the isomorphism class of $\qcs{L}$, so we obtain a map
\begin{align*}
\TrFrob{G} : \QCiso{G} &\to G(\Fq)^* \\
\qcs{L} &\mapsto \trFrob{\qcs{L}}.
\end{align*}
Since tensor products on the stalks of $\qcs{L}$ induce pointwise multiplication on the trace of Frobenius, $\TrFrob{G}$ is a group homomorphism.  We devote the remainder of this section to
studying the behavior of the trace of Frobenius under morphisms of group schemes.  These simple but important
results on pullbacks and external products will be used in the proofs of many results in this paper.

\begin{lemma}\label{lem:pullback}
  If $f : H\to G$ is a morphism of smooth commutative group schemes over $\Fq$ then
  \begin{align*}
  f^* : \QC(G) &\to \QC(H) \\
  (\gqcs{L},\mu,\phi) &\mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)
  \end{align*}
  defines a monoidal functor dual to $f \colon H(\Fq) \to G(\Fq)$:
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\TrFrob{G}} & \QCiso{H} \dar{\TrFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{lemma}
\begin{proof}
  Let $\qcs{L}$ be a quasicharacter sheaf on $G$. 
  Pullback by $\bf$ takes rank~one local systems to rank~one local systems.
  To see that $(\bfxf)^* \mu$ satisfies \ref{QC.2},
  apply the functor $(\bfxf)^*$
  to \ref{QC.2} for $\qcs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gqcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gqcs{L}$.
  To show that $f^*\qcs{L}$ satisfies
  \ref{QC.3}, apply the same functor to \ref{QC.3} for $\qcs{L}$.
  Since $f$ is a morphism of group schemes defined over $\Fq$
  it provides isomorphisms $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG}^* (\bfxf)^*$
  and $(\bfxf)^* \bm^*\iso \bm^* \bf^*$ between functors of constructible sheaves.

@@

Now, let $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
Set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$;
by Lemma~\ref{lem:pullback}, $\qcs{L}_1$ a quasicharacter sheaf on $G_1$
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
Applying the functor $f^*$ to the isomorphism $\mu$ yields
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already seen that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$.  
Since $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L} = r_1^* \gqcs{L}_1$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L} = r_2^*\gqcs{L}_2$,
we have 
\[
f^*(\gqcs{L}\boxtimes \gqcs{L})  = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L} = \gqcs{L}_1\boxtimes \gqcs{L}_2.
\]
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies
\ref{QC.3} as it applies here,
from which it follows that we have exhibited an isomorphism
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

\begin{proposition}\label{prop:functorialG}
Trace of Frobenius $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is a natural transformation
between the two contravariant additive functors
\begin{align*}
F_1 : G &\mapsto \QCiso{G} \\
F_2 : G &\mapsto G(\Fq)^*
\end{align*}
from the category of smooth commutative group schemes over $\Fq$ to the category of abelian groups.
%Likewise, for $\TrFrob{G} : \lQCiso{G} \to G(\Fq)^*$, {\it mut.\hskip-1pt mut.}
\end{proposition}

@@

The equivalence $G \mapsto G(\bFq)$
from the category of \'etale commutative group schemes over $\Fq$ to the category of commutative groups equipped
with an action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on the group,
provides the following simple description of quasicharacter sheaves.
%
A quasicharacter sheaf $\qcs{L}$ on an \'etale commutative group scheme $G$ over $\Fq$ is:
\begin{enumerate}
 \labitem{(qc.1)}{qc.1} an indexed set of one-dimensional
  $\EE$-vector spaces $\gqcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

 \labitem{(qc.2)}{qc.2} an indexed set of isomorphisms
  $\mu_{x,y} : \gqcs{L}_{x+y} \xrightarrow{\iso} \gqcs{L}_{x} \otimes\gqcs{L}_{y}$,
  for all $x,y \in G(\bFq)$, such that
  \[
   \begin{tikzcd}[row sep=40]
    \gqcs{L}_{x+y+z} \arrow{rr}{\mu_{x+y,z}} \arrow[swap]{d}{\mu_{x,y+z}}
    && \gqcs{L}_{x+y}\otimes\gqcs{L}_{z} \dar{\mu_{x,y} \tight{0.5}{\otimes}{1} \id} \\
    \gqcs{L}_{x} \otimes\gqcs{L}_{y+z} \arrow{rr}{\id \otimes\mu_{y,z}}
    && \gqcs{L}_{x} \otimes\gqcs{L}_{y} \otimes\gqcs{L}_{z}
   \end{tikzcd}
  \]
  commutes, for all $x,y,z\in G(\bFq)$; and
 \labitem{(qc.3)}{qc.3} an indexed set of isomorphisms $\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$
  such that
  \[
   \begin{tikzcd}[row sep=40]
    \gqcs{L}_{\Frob{}(x)+\Frob{}(y)} \arrow[swap]{d}{\phi_{x+y}} \arrow{rr}{\mu_{\Frob{}(x),\Frob{}(y)}}
    && \gqcs{L}_{\Frob{}(x)}\otimes\gqcs{L}_{\Frob{}(y)} \dar{\phi_x \tight{0}{\otimes}{0} \phi_y} \\
    \gqcs{L}_{x+y} \arrow{rr}{\mu_{x,y}}
    && \gqcs{L}_x \otimes\gqcs{L}_y
   \end{tikzcd}
  \]
  commutes, for all $x,y\in G(\bFq)$.
\end{enumerate}
Under this equivalence, a morphism $\rho : \qcs{L} \to \qcs{L'}$ of quasicharacter sheaves on $G$ is given by
\begin{enumerate}
 \labitem{(qc.4)}{qc.4} an indexed set $\rho_x : \gqcs{L}_x \to \gqcs{L'}_x$
  of linear transformations such that
  \[
   \begin{tikzcd}[column sep=40]
    \arrow[swap]{d}{\phi_x} \gqcs{L}_{\Frob{}(x)} \rar{\rho_{\Frob{}(x)}} & \gqcs{L'}_{\Frob{}(x)} \dar{\phi_x'}
    &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}
    & \arrow[swap]{d}{\mu_{x,y}} \gqcs{L}_{x+y} \rar{\rho_{x+y}} & \gqcs{L'}_{x+y} \dar{\mu'_{x,y}} \\
    \gqcs{L}_x \rar{\rho_x} & \gqcs{L'}_x
    & {} & \gqcs{L}_x\otimes\gqcs{L}_y \rar{\rho_x\otimes\rho_y} & \gqcs{L'}_x \otimes\gqcs{L'}_y
   \end{tikzcd}
  \]
  both commute for all $x, y \in G(\bFq)$.
\end{enumerate}

@@

The zeroth page of the Hochschild-Serre spectral sequence, as in \cite{Vakil:spectral_sequences} following \cite{Weibel:Homological_Algebra}*{Ch 5 and \S 7.5}, for example,
is a double complex $E^{\bullet, \bullet}$ defined by
\[
E^{i,j} = C^i(\Weil{}, C^j(G(\bFq), \EEx)),
\]
The derivative of group cohomology yields two derivatives:
\begin{align*}
d_0 = d_G &: E^{i,j} \to E^{i,j+1} \\
d_{\Weil{}} &: E^{i,j} \to E^{i+1,j}.
\end{align*}
Combining these yields a derivative $d = d_G + (-1)^j d_{\Weil{}}$ on the total complex
\[
E^n_G = \bigoplus_{i+j=n} E^{i,j}.
\]
The machinery of spectral sequences gives us a sequence of pages $E_r^{i,j}$, converging to a page $E_{\infty}^{i,j}$. We summarize the key properties of this spectral sequence in the following proposition.

\begin{proposition}
\begin{enumerate}
\item The second page is given by $E_2^{i,j} = \Hh^i(\Weil{}, \Hh^j(G(\bFq), \EEx))$,
\item there is an isomorphism $\Hh^n(\Weil{} \ltimes G(\bFq), \EEx) \cong \Hh^n(E_G^\bullet)$,
\item there is a filtration $\Hh^n(\Weil{} \ltimes G(\bFq), \EEx) = F_n \supset \cdots \supset F_0 \supset F_{-1} = 0$ where $F_i / F_{i-1} \cong E_{\infty}^{i, n-i}$.
\end{enumerate}
\end{proposition}
%\todo{Is this a lemma? Is the corollary a proposition?}

@@

\begin{corollary}
There is a short exact sequence
 \[
  \begin{tikzcd}
    0 \arrow{r}
    & \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \arrow{r}
    & \Hh^2(E^\bullet_G) \arrow{r}
    & \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \arrow{r}
    & 0
  \end{tikzcd}
 \]
\end{corollary}

@@

In this section we define an isomorphism
\[
S_G : \QCiso{G} \to \Hh^2(E^\bullet_G).
\]
Let $\qcs{L} = (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
For each geometric point $x\in {\bar G}$, choose a basis $\{ v_x \}$ for $\gqcs{L}_x$.
Through this choice, $\qcs{L}$ determines functions
\begin{align*}
a : {\bar G}\times {\bar G} &\to \EEx & b : {\bar G} &\to \EEx \\
\mu_{x,y}(v_{x+y}) &= a(x,y) v_x \otimes v_y & \phi_x(v_{\Frob{G}(x)}) &= b(x) v_x.
\end{align*}
Condition~\ref{QC.2} implies that
\begin{equation}\label{2-cocyle}
a(x+y,z) a(x,y) = a(x,y+z) a(y,z)
\end{equation}
for all $x,y,z\in {\bar G}$, so $a \in Z^2({\bar G},\EEx)$.  Similarly, condition~\ref{QC.3} gives
\begin{equation}\label{nohom}
\frac{a(\Frob{G}(x),\Frob{G}(y))}{a(x,y)} =  \frac{b(x+y)}{b(x) b(y)}
\end{equation}
for $x, y \in {\bar G}$.
Let $\alpha \in C^0(\Weil{},C^2({\bar G},\EEx)$ be the $0$-cochain corresponding to $a$ and let $\beta\in C^1(\Weil{},C^1({\bar G},\EEx)$ be the cocycle such that $\beta(\Frob{})$ is $b$ \emph{written additively}.
Then
\[
d_G\alpha =0, \qquad\qquad
d_{\Weil{}} \alpha = d_{G} \beta,\qquad\qquad
d_{\Weil{}} \beta =0;
\]
in other words,
\[\alpha\oplus \beta \in Z^2(E^\bullet_G).\]
Although the cocycle $\alpha\oplus \beta$ is not well defined by $\qcs{L}$, its class in $Z^2(E^\bullet_G)$ is.
To see this, let $\{ v'_x \in \gqcs{L}_x^\times \tq x \in {\bar G}\}$ be another choice and let $\alpha'\oplus \beta' \in Z^2(E^\bullet_G)$ be defined by $\qcs{L}$ and this choice, as above.
Now let $\delta \in C^0(\Weil{},C^1({\bar G},\EEx))$ be the additive version of the function $d : {\bar G}\to \EEx$ defined by $v'_x = d(x) v_x$.
Then, chasing through \ref{QC.2} and \ref{QC.3} we find
\[
\alpha'\oplus\beta' = \alpha\oplus\beta + d\delta,
\]
so the class $[\alpha\oplus\beta]$ of $\alpha\oplus\beta$ in $\Hh^2(E^\bullet_G)$ is independent of the choice made above. It is also easy to see that $[\alpha\oplus\beta] = [\alpha_0\oplus\beta_0]$ when $\qcs{L} \iso \qcs{L}_0$,
which concludes the definition of the function
\begin{align*}
S_G : \QCiso{G} &\to \Hh^2(E^\bullet_G)\\
[\qcs{L}] &\mapsto [\alpha\oplus \beta].
\end{align*}
It is also easy to see that $[\alpha_1\oplus\beta_1] + [\alpha_2\oplus\beta_2] = [\alpha_3\oplus\beta_3]$ when $\qcs{L}_3 = \qcs{L}_1\otimes \qcs{L}_2$, so $S_G$ is a group homomorphism.

@@

\begin{proposition}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
%We first show that $S_G$ is injective.  
Suppose $\qcs{L} \in \QCiso{G}$ with $S_G(\qcs{L}) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\sigma$ for some $\sigma \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\sigma_x : \gqcs{L}_x \to \EEx$ by $v_x \mapsto \sigma(x)$.
Then the indexed set of isomorphisms $\{ \sigma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

@@

\begin{definition}
Let $\bQC(G)$ be the category of pairs $(\qcs{E},\mu_\qcs{E})$
where $\qcs{E}$ an $\ell$-adic local system on $G$ of rank~one,
equipped with an isomorphism $\mu_\qcs{E} : m^* \qcs{E} \to \qcs{E} \boxtimes \qcs{E}$
satisfying the analogue of \ref{QC.1} on $G$;
morphisms in $\bQC(G)$ are defined as in the second part of
\ref{QC.3}.
\end{definition}

>>>>

\begin{abstract}
By extending the scope of function-sheaf dictionary for connected commutative algebraic groups, we find a geometrization of characters of all smooth commutative group schemes over finite fields.
As an application, we explain how to geometrize and categorify quasicharacters of all connected commutative algebraic groups over non-Archimedean local fields that admit a locally finite type Neron model, such as algebraic tori, abelian varieties and unipotent K-wound groups.
\end{abstract}

@@

We will describe the group $\QCiso{G}$ in Theorem~\ref{thm:snake}
and the sets $\Hom(\qcs{L},\qcs{L}')$ in Theorem~\ref{thm:autornaught}; in this way we provide a complete description of the category $\QC(G)$.
%
In the meantime, we close this section with an elementary observation about $\Hom(\qcs{L},\qcs{L}')$.
We call a morphism of sheaves \emph{trivial} if it is zero on every stalk.

\begin{lemma}\label{lem:autornaught}
Let $G$ be a smooth commutative group scheme over $\Fq$.
If $\qcs{L}$ and $\qcs{L}'$ are quasicharacter sheaves on $G$, then
every $\rho\in \Hom(\qcs{L},\qcs{L}')$ is either trivial or an isomorphism. 
\end{lemma}

@@

Now suppose $\qcs{L} = (\gqcs{L},\mu,\phi)$ is a quasicharacter sheaf on $G$.
Then the isomorphism $\bm^* \gqcs{L} \iso \gqcs{L} \boxtimes\gqcs{L}$ and the diagram of
\ref{QC.3} guarantee
that the function $\trFrob{(\gqcs{L},\phi)} : G(\Fq)\to \EEx$ is a group homomorphism, which we will also denote by $\trFrob{\qcs{L}}$.  Moreover,
this homomorphism depends only on the isomorphism class of $\qcs{L}$, so we obtain a map
\begin{align*}
\TrFrob{G} : \QCiso{G} &\to G(\Fq)^*, \\
\qcs{L} &\mapsto \trFrob{\qcs{L}}.
\end{align*}
Since tensor products on the stalks of $\qcs{L}$ induce pointwise multiplication on the trace of Frobenius, $\TrFrob{G}$ is a group homomorphism.  We devote the remainder of this section to
studying the behavior of the trace of Frobenius under morphisms of group schemes.  These simple but important
properties of pullbacks and external products will be used in the proofs of many results in this paper.

\begin{lemma}\label{lem:pullback}
  If $f : H\to G$ is a morphism of smooth commutative group schemes over $\Fq$, then
  \begin{align*}
  f^* : \QC(G) &\to \QC(H) \\
  (\gqcs{L},\mu,\phi) &\mapsto (\bf^*\gqcs{L},(\bfxf)^*\mu,\bf^*F)
  \end{align*}
  defines a monoidal functor dual to $f \colon H(\Fq) \to G(\Fq)$ in the sense that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \QCiso{G} \rar{f^*} \arrow[swap]{d}{\TrFrob{G}} & \QCiso{H} \dar{\TrFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{lemma}
\begin{proof}
  Let $\qcs{L}$ be a quasicharacter sheaf on $G$. 
  Pullback by $\bf$ takes rank-one local systems to rank-one local systems.
  To see that $(\bfxf)^* \mu$ satisfies \ref{QC.2},
  apply the functor $(\bfxf)^*$
  to \ref{QC.2} for $\qcs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gqcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gqcs{L}$.
  To show that $f^*\qcs{L}$ satisfies
  \ref{QC.3}, apply the same functor to \ref{QC.3} for $\qcs{L}$.
  Since $f$ is a morphism of group schemes defined over $\Fq$
  it provides isomorphisms $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG}^* (\bfxf)^*$
  and $(\bfxf)^* \bm^*\iso \bm^* \bf^*$ between functors of constructible sheaves.

@@

Now, let $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$
and set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$.
By Lemma~\ref{lem:pullback}, $\qcs{L}_1$ is a quasicharacter sheaf on $G_1$
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
Applying the functor $f^*$ to the isomorphism $\mu$ yields
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already seen that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$.  
Since $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L} = r_1^* \gqcs{L}_1$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L} = r_2^*\gqcs{L}_2$,
we have 
\[
f^*(\gqcs{L}\boxtimes \gqcs{L})  = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L} = \gqcs{L}_1\boxtimes \gqcs{L}_2.
\]
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies
\ref{QC.4} as it applies here,
from which it follows that we have exhibited an isomorphism
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

\begin{proposition}\label{prop:functorialG}
The homomorphism $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ defines a natural transformation
between the two contravariant additive functors
\begin{align*}
F_1 : G &\mapsto \QCiso{G} \\
F_2 : G &\mapsto G(\Fq)^*
\end{align*}
from the category of smooth commutative group schemes over $\Fq$ to the category of abelian groups.
%Likewise, for $\TrFrob{G} : \lQCiso{G} \to G(\Fq)^*$, {\it mut.\hskip-1pt mut.}
\end{proposition}

@@

The equivalence $G \mapsto G(\bFq)$,
from the category of \'etale commutative group schemes over $\Fq$ to the category of commutative groups equipped
with a continuous action of $\Gal(\bFq/\Fq)$,
provides the following simple description of quasicharacter sheaves.
%
A quasicharacter sheaf $\qcs{L}$ on an \'etale commutative group scheme $G$ over $\Fq$ is:
\begin{enumerate}
 \labitem{(qc.1)}{qc.1} an indexed set of one-dimensional
  $\EE$-vector spaces $\gqcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

 \labitem{(qc.2)}{qc.2} an indexed set of isomorphisms
  $\mu_{x,y} : \gqcs{L}_{x+y} \xrightarrow{\iso} \gqcs{L}_{x} \otimes\gqcs{L}_{y}$,
  for all $x,y \in G(\bFq)$, such that
  \[
   \begin{tikzcd}[row sep=40]
    \gqcs{L}_{x+y+z} \arrow{rr}{\mu_{x+y,z}} \arrow[swap]{d}{\mu_{x,y+z}}
    && \gqcs{L}_{x+y}\otimes\gqcs{L}_{z} \dar{\mu_{x,y} \tight{0.5}{\otimes}{1} \id} \\
    \gqcs{L}_{x} \otimes\gqcs{L}_{y+z} \arrow{rr}{\id \otimes\mu_{y,z}}
    && \gqcs{L}_{x} \otimes\gqcs{L}_{y} \otimes\gqcs{L}_{z}
   \end{tikzcd}
  \]
  commutes, for all $x,y,z\in G(\bFq)$; and
 \labitem{(qc.3)}{qc.3} an indexed set of isomorphisms $\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$
  such that
  \[
   \begin{tikzcd}[row sep=40]
    \gqcs{L}_{\Frob{}(x)+\Frob{}(y)} \arrow[swap]{d}{\phi_{x+y}} \arrow{rr}{\mu_{\Frob{}(x),\Frob{}(y)}}
    && \gqcs{L}_{\Frob{}(x)}\otimes\gqcs{L}_{\Frob{}(y)} \dar{\phi_x \tight{0}{\otimes}{0} \phi_y} \\
    \gqcs{L}_{x+y} \arrow{rr}{\mu_{x,y}}
    && \gqcs{L}_x \otimes\gqcs{L}_y
   \end{tikzcd}
  \]
  commutes, for all $x,y\in G(\bFq)$.
\end{enumerate}
Under this equivalence, a morphism $\rho : \qcs{L} \to \qcs{L'}$ of quasicharacter sheaves on $G$ is given by
\begin{enumerate}
 \labitem{(qc.4)}{qc.4} an indexed set $\rho_x : \gqcs{L}_x \to \gqcs{L'}_x$
  of linear transformations such that
  \[
   \begin{tikzcd}[column sep=40]
    \arrow[swap]{d}{\phi_x} \gqcs{L}_{\Frob{}(x)} \rar{\rho_{\Frob{}(x)}} & \gqcs{L'}_{\Frob{}(x)} \dar{\phi_x'}
    &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}
    & \arrow[swap]{d}{\mu_{x,y}} \gqcs{L}_{x+y} \rar{\rho_{x+y}} & \gqcs{L'}_{x+y} \dar{\mu'_{x,y}} \\
    \gqcs{L}_x \rar{\rho_x} & \gqcs{L'}_x
    & {} & \gqcs{L}_x\otimes\gqcs{L}_y \rar{\rho_x\otimes\rho_y} & \gqcs{L'}_x \otimes\gqcs{L'}_y
   \end{tikzcd}
  \]
  both commute, for all $x, y \in G(\bFq)$.
\end{enumerate}

@@

The zeroth page of the Hochschild-Serre spectral sequence
is a double complex $E^{\bullet, \bullet}$ defined by
\[
E^{i,j} = C^i(\Weil{}, C^j(G(\bFq), \EEx));
\]
see \cite{Vakil:spectral_sequences}, which follows \cite{Weibel:Homological_Algebra}*{Ch 5 and \S 7.5}.
The standard derivative on cochains yields two derivatives,
\begin{align*}
d_G &: E^{i,j} \to E^{i,j+1} \mbox{and} \\
d_{\Weil{}} &: E^{i,j} \to E^{i+1,j};
\end{align*}
we use the first as the derivative $d_0$ and the second to induce $d_1$.
Combining them also yields a derivative $d = d_G + (-1)^j d_{\Weil{}}$ on the total complex
\[
E^n_G = \bigoplus_{i+j=n} E^{i,j}.
\]
The machinery of spectral sequences gives us a sequence of pages $E_r^{i,j}$, converging to a page $E_{\infty}^{i,j}$. We summarize the key properties of this spectral sequence in the following proposition.

\begin{proposition} In the spectral sequence defined above,
\begin{enumerate}
\item the second page is given by $E_2^{i,j} = \Hh^i(\Weil{}, \Hh^j(G(\bFq), \EEx))$,
\item there is an isomorphism $\Hh^n(\Weil{} \ltimes G(\bFq), \EEx) \cong \Hh^n(E_G^\bullet)$, and
\item there is a filtration $\Hh^n(\Weil{} \ltimes G(\bFq), \EEx) = F_n \supset \cdots \supset F_{-1} = 0$, with $F_i / F_{i-1} \cong E_{\infty}^{i, n-i}$.
\end{enumerate}
\end{proposition}
%\todo{Is this a lemma? Is the corollary a proposition?}

@@

\begin{corollary}
There is a short exact sequence
 \[
    0 \to
    \Hh^0(\Weil{},\Hh^2({\bar G},\EEx)) \to
    \Hh^2(E^\bullet_G) \to
    \Hh^1(\Weil{},\Hh^1({\bar G},\EEx)) \to
    0
 \]
\end{corollary}

@@

In this section we define an isomorphism
\[
S_G : \QCiso{G} \to \Hh^2(E^\bullet_G).
\]
Let $\qcs{L} = (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$.
For each geometric point $x\in {\bar G}$, choose a basis $\{ v_x \}$ for $\gqcs{L}_x$.
Through this choice, $\qcs{L}$ determines functions
\begin{align*}
a : {\bar G}\times {\bar G} &\to \EEx & b : {\bar G} &\to \EEx \\
\mu_{x,y}(v_{x+y}) &= a(x,y) v_x \otimes v_y & \phi_x(v_{\Frob{G}(x)}) &= b(x) v_x.
\end{align*}
Condition~\ref{QC.2} implies that
\begin{equation}\label{2-cocyle}
a(x+y,z) a(x,y) = a(x,y+z) a(y,z)
\end{equation}
for all $x,y,z\in {\bar G}$, so $a \in Z^2({\bar G},\EEx)$.  Similarly, condition~\ref{QC.3} gives
\begin{equation}\label{nohom}
\frac{a(\Frob{G}(x),\Frob{G}(y))}{a(x,y)} =  \frac{b(x+y)}{b(x) b(y)}
\end{equation}
for $x, y \in {\bar G}$.
Let $\alpha \in C^0(\Weil{},C^2({\bar G},\EEx)$ be the $0$-cochain corresponding to $a$ and let $\beta\in C^1(\Weil{},C^1({\bar G},\EEx)$ be the cocycle such that $\beta(\Frob{})$ is $b$.  We will write both $\alpha$ and $\beta$ additively.
Then
\[
d_G\alpha =0, \qquad\qquad
d_{\Weil{}} \alpha = d_{G} \beta,\qquad\qquad
d_{\Weil{}} \beta =0;
\]
in other words,
\[\alpha\oplus \beta \in Z^2(E^\bullet_G).\]
Although the cocycle $\alpha\oplus \beta$ is not well defined by $\qcs{L}$, its class in $Z^2(E^\bullet_G)$ is.
To see this, let $\{ v'_x \in \gqcs{L}_x^\times \tq x \in {\bar G}\}$ be another choice and let $\alpha'\oplus \beta' \in Z^2(E^\bullet_G)$ be defined by $\qcs{L}$ and this choice, as above.
Now let $\delta \in C^0(\Weil{},C^1({\bar G},\EEx))$ correspond to the function $d : {\bar G}\to \EEx$ defined by $v'_x = d(x) v_x$.
Chasing through \ref{QC.2} and \ref{QC.3}, we find
\[
\alpha'\oplus\beta' = \alpha\oplus\beta + d\delta,
\]
so the class $[\alpha\oplus\beta]$ of $\alpha\oplus\beta$ in $\Hh^2(E^\bullet_G)$ is independent of the choice made above. It is also easy to see that $[\alpha\oplus\beta] = [\alpha_0\oplus\beta_0]$ when $\qcs{L} \iso \qcs{L}_0$,
which concludes the definition of the function
\begin{align*}
S_G : \QCiso{G} &\to \Hh^2(E^\bullet_G)\\
[\qcs{L}] &\mapsto [\alpha\oplus \beta].
\end{align*}
It is also easy to see that $[\alpha_1\oplus\beta_1] + [\alpha_2\oplus\beta_2] = [\alpha_3\oplus\beta_3]$ when $\qcs{L}_3 = \qcs{L}_1\otimes \qcs{L}_2$, so $S_G$ is a group homomorphism.

@@

\begin{proposition}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
%We first show that $S_G$ is injective.  
Suppose $[\qcs{L}] \in \QCiso{G}$ with $S_G([\qcs{L}]) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\sigma$ for some $\sigma \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\sigma_x : \gqcs{L}_x \to \EEx$ by $v_x \mapsto \sigma(x)$.
Then the indexed set of isomorphisms $\{ \sigma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

@@

\begin{definition}
Let $\bQC(G)$ be the category of pairs $(\qcs{E},\mu_\qcs{E})$
where $\qcs{E}$ an $\ell$-adic local system on $G$ of rank-one,
equipped with an isomorphism $\mu_\qcs{E} : m^* \qcs{E} \to \qcs{E} \boxtimes \qcs{E}$
satisfying the analogue of \ref{QC.1} on $G$;
morphisms in $\bQC(G)$ are defined as in the second part of
\ref{QC.3}.
\end{definition}
